---
title: "Project - Cardio Good Fitness"
author: "Adrian Calvin"
output:
  pdf_document: 
    toc: yes
    number_sections: yes
  html_notebook::
    toc: true
---


\pagebreak
# Project Objective
The objective of this report is to explore the Cardio Machine dataset and gain 
insights to assist in business decisions. This exploration report will consist 
of the following:

* Importing the dataset in R
* Identifying customer profiles
* Understanding the structure of dataset
* Exploratory data analysis using Graphical exploration and Descriptive statistics

# Assumptions
The data provided at least somewhat realistically models a simple exercise-machine 
business in the early 21st century (i.e. people are more fitness minded, the
internet exists, et. al.)

In addition, it is assumed that the higher the model number, the higher-end the 
product, and the higher the price.

We will assume the sample is taken from American customers.

For explanatory purposes, we will assign the following prices to each product:

```{r}
prices <- list()
prices['TM195'] <- 10
prices['TM498'] <- 15
prices['TM798'] <- 20
prices.table <- data.frame(
  matrix(c("TM195", 
           "TM498", 
           "TM798", 
           prices['TM195'], 
           prices['TM498'], 
           prices['TM798']), nrow = 3))
names(prices.table) <- c("Product", "Price")
prices.table
```


# Data Import
```{r}
# Import Libraries
suppressMessages(library(lubridate))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(scales))
suppressMessages(library(gridExtra))
suppressMessages(library(corrplot))
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))

# Import Data
cardio <- read.csv("CardioGoodFitness.csv")
```
\pagebreak

# First Impressions of Data
## Variable Identification
The purpose of this section is to get basic familiarity with the data set's character.

### Dimensions

```{r}
dim(cardio)
```

### Head
```{r}
head(cardio)
```

### Summary
```{r}
summary(cardio)
```

**Key Take-away**

<p>
_This sample appears to be substantially better educated the 
general population_.
</p>v

In 2018, according to the U..S Census Bureau, only 34.98% of the U.S. population
 (25 and older) is estimated to have attained a bachelor's degree, or higher.  The mean education
level of the sample is 15.57, indicating that on average, a skew towards higher
education.  

This makes sense, as the "common sense" method would tell a person that more 
educated people are more likely to be sensitive to new knowledge that can better
their well-being.  That being said, I cannot support this 
conclusion with the data given.

```{r, echo=FALSE}

```
1. https://en.wikipedia.org/wiki/Educational_attainment_in_the_United_States#General_attainment_of_degrees/diplomas
\pagebreak
 
# Data Mangling
 
```{r}
 # Add columns for grouping Income, Age, and Education
cardio <- mutate(
  cardio, 
  IncomeGroup = cut(
     Income, 
     breaks   = c(0, 20000-1,30000-1,45000-1,60000-1,100000-1,Inf), 
     labels   = c("[0-19k)", "[20-30k)", "[30-45k)", "[45-60k)", "[60k-100k)", ">= 100k")),
  
  AgeGroup    = cut(Age, breaks = c(17, 19, 25, 29, 36, 59)),
  
  EducGroup   = cut(Education, breaks = c(0,12,15,16,99), 
     labels   = c("High School", "Some College", "Bachelors", "(Some)Grad School")))

# Factorize other columns
cardio$Gender = as.factor(cardio$Gender)
cardio$Education     = as.factor(cardio$Education)
cardio$MaritalStatus = as.factor(cardio$Gender)
cardio$Product       = as.factor(cardio$Product)

head(cardio)
```
\pagebreak

# Customer Profiles
Below is a break-down of purchases by three criteria: Education, Age, Income, and
self-perceived fitness level groupings:

 

## Product Sales by Group

### Education

```{r}
ggplot(cardio, aes(x   = EducGroup, fill = Product)) + 
            labs(title = "Product by Education Group", 
                     x = "Education Group", 
                     y = "Product Count") + 
     geom_bar(position = "stack")
```
_As education level increases, so do proportion of the more expensive model in 
terms of sales._

\pagebreak

### Age

```{r}
ggplot(cardio, aes(x  = AgeGroup, fill = Product)) + 
           labs(title = "Product by Age Group", 
                    x = "Age Group", 
                    y = "Product Count") + 
    geom_bar(position = "stack")
```
_Unsurprisingly, the 19 - 25 age group invests in the most purchases.  Somewhat surprisingly,
however, the ratio of the mid-range TM498 to high-end TM798s did not have a 
steady decline with increasing age._

\pagebreak

### Income

```{r}
ggplot(cardio,  aes(x = IncomeGroup, fill = Product)) + 
          labs(title  = "Product by Income Group", 
                   x  = "Income Group", 
                   y  = "Product Count") + 
    geom_bar(position = "stack")
```
_Predictably, the more expensive model rises in popularity, as a percentage of total sales, as income level rises._

\pagebreak

### Fitness 

```{r}
ggplot(cardio, aes(x   = Fitness, fill = Product)) + 
            labs(title = "Product by Fitness Level", 
                     x = "Fitness Level", 
                     y = "Product Count") + 
     geom_bar(position = "stack")
```
_This subjective rating of fitness indicates that the more fit a person believed
their self to be, the more they were willing to pay - at least at the higher levels.
The fitness level does not seem to have a huge impact on the choice of product at
average level (3) and below._

\pagebreak
 
